Brownian motion of main-belt asteroids on human timescales

### 2025-03-02 main.tex
Edit run-through done, just need equation check.

\newcommand{\Var}{\textrm{Var}}
\newcommand{\vcirc}{v_c}
\newcommand{\vrel}{v_{\rm r}}
\newcommand{\Msun}{M_\odot}
\newcommand{\covm}{C}
\newcommand{\lop}{\varpi}   % longitude of perihelion
\newcommand{\ma}{\ell}  % Mean anomaly

@@

High-precision measurements of positions of Solar System bodies have
great value to multiple scientific pursuits.  They can reveal
gravitational accelerations due to undiscovered bodies such as a
hypothetical ``Planet X'' \citep[\eg][]{holmanP9,inpopP9,trojans,occultations} or passing primordial
black holes \citep[\eg][]{pbh}, determine
the masses of known bodies \citep[\eg][]{goffin,baer}, test the laws
of physics \citep[\eg][]{inpopgraviton}, and characterize 
the non-gravitational forces that drive many forms of planetary
migration, \eg\ the maintainence of the near-Earth asteroid (NEA)
population \citep[reviewed by][]{yarkovsky}.  While most applications of precision orbit determination
have used the 8 major planets as test particles, there are now $>10^6$
known minor planets available for this task.  These small bodies' astrometric
information content is growing extremely rapidly through large-scale
survey projects that can produce
milliarcsecond-scale accuracy for magnitude-limited small-body populations, such as Gaia \citep{gaiass3,gaiafpr},
PanSTARRS,\footnote{\url{https://www2.ifa.hawaii.edu/research/Pan-STARRS.shtml}}
and particularly the upcoming Rubin observatory
\textit{Legacy Survey of Space and Time} (LSST).\footnote{\url{https://rubinobservatory.org/}}  The great majority
of currently-tracked objects are main-belt asteroids (MBAs), so we
wish to investigate the potential power of MBAs as gravitational test
bodies.

@@

In principal, one can collect all of the positional information for
all $>10^6$ known solar-system bodies, and fit them to a model with
a free state vector and mass for each body.  Calculating the influence
of $>7\times10^6$ free parameters on $10^8$ or so observables is,
however, unlikely to be computationally feasible, and most of the
bodies' masses will be lost in measurement noise and/or be degenerate
with other bodies' masses.  The most ambitious attempt at a global
solution that we are aware of is that of \citet{goffin}, who was able
to include 250 MBAs as gravitating bodies in the fit to the full
corpus of observations at the time.  For only $\approx 1/2$ of the
bodies did the fit yield masses $>3\times$ the uncertainties.  With
vast improvements in data and computing power since then and in the
next decade, the number of detectable
asteroid masses should increase, but there will always be
uncertainties that accrue from the unmodelled deflectors and the
finite $\sigma_M$ of modelled deflectors.  These remnant uncertainties
will in turn serve as noise in attempts to model new sources of
acceleration.

@@

\section{Calculation overview}
We aim to calculate the Brownian motion variance to an accuracy of a
factor $\approx 2.$
Greater accuracy in the calculation is not warranted
since some of the principal inputs have substantial uncertainty.  The
number of MBAs vs mass is 
uncertain because of unknowns in the conversion
from $H$ to mass; the status of future observations and $\sigma_M$ is
also unclear.  We are hence justified in retaining only the
contributions to our results that are leading
order in the MBA orbital eccentricity $e$ of the tracer population.
We will assume throughout that the joint
distribution of mass (or $H$) and orbital elements is separable into a
mass distribution and an orbital distribution.  In this scenario, the
Brownian motion is independent of the tracer mass, and involves an
integral over the mass distribution of the deflecting bodies.

We adopt a cylindrical coordinate system $(r,\phi,z),$ with $\zhat$
normal to the initial orbital plane, and $\phi=z=0$ toward the
perhelion of the initial orbit, \ie\ both the initial ascending node
$\Omega_0$ and longitude of perihelion $\lop_0$ equal to zero.  The
unit vectors $\rhat, \phat$ rotate with the target asteroid.  The mean
anomaly at time $t=0$ is $\ma_0.$

@@

We describe all encounters with other asteroids in the impulse approximation, defining $\vecI$ as the $\Delta\vecv$ imparted on the target by the deflector.
In our units, the gravitational impulse imparted by a deflector of mass $M_d$ approaching at impact parameter $\vecb=b\bhat$ and relative velocity $\vrel$ is
\begin{equation}
  \vecI = 2 \frac{M_d}{\Msun} (bv)^{-1} \bhat.
  \label{eq:impulse}
\end{equation}
All of our results will be derived at leading order in $\vecI,$ which
is very well justified by the small size of MBA-induced impulses.

@@

From this knowledge of the impulse distribution, our goal is to obtain
the covariance matrix of the deviations of the target's position from
the initial orbit, after some time $T.$ The position observables are
the range, plus the celestial latitude and longitude of the target.  We will simplify our results by assuming a heliocentric observer, so that the observational position vector is $\vecp\equiv (r,\phi,\theta=z/r).$ The quantity we seek is the covariance matrix $\covm^p$ of the observations attributable to the accumulated gravitational perturbations:
\begin{equation}
  \covm^p  \equiv  \left\langle  \Delta\vecp \Delta\vecp^T\right\rangle,  \label{eq:Cp}
\end{equation}
where the angle brackets indicate an average over possible
realizations of the impulse history, and over the mean anomaly $\ma$ at the time of observation.  To do so, we will introduce an intermediate set of 6 parameters $\vecq$ describing the orbital elements induced by the impulses.  The $\vecq$ components will be selected to respond linearly to $\vecI$ at $|I|\ll 1.$
We will derive the matrix $\matA$ that describes the orbital-element shifts at time $T$ that arise from an impulse at time $t_i$: 
\begin{equation}
  \Delta \vecq(e_0,T,t_i) = \matA(e_0, T, t_i) \, \vecI.
  \label{eq:A}
\end{equation}
In our first-order perturbation theory, $\Delta\vecq(T)$ will be the sum of the $\Delta\vecq_i$ imparted by all impulses applied at times $0<t_i<T.$  Because the impulses are uncorrelated, $\vecq$ will therefore be the result of a random walk.  The distribution will have a covariance matrix $\covm^q(T) \equiv \left\langle \Delta\vecq(T) \Delta\vecq^T(T) \right\rangle$ whose elements are
\begin{eqnarray}
  \covm^q_{jk}(e_o,T) & = & \left\langle \sum_{i,\gamma} \matA_{j\gamma}(e_0,T,t_i) \matA_{k\gamma}(e_0,T,t_i) I^2_{i,\gamma} \right\rangle \\
           & = & \sum_\gamma \int dt_i \matA_{j\gamma}(e_0,T,t_i) \matA_{k\gamma}(e_0,T,t_i) \left\langle n I_\gamma^2\right\rangle.
\label{eq:Cqjk}
\end{eqnarray}
In the first line, the sum $i$ runs over the impulses and $\gamma$ runs over the components $r,\phi,z$ of the impulse. The second line evaluates the expectation value of averaging over realizations of the random walk of impulses, exploiting the independence of the individual impulses from each other.

The last element of our calculation will be a conversion from the
element shifts $\Delta\vecq$ into the observed displacements $\Delta\vecp$
at time $T.$  In linear perturbation theory this will again be expressible as a matrix
\begin{eqnarray}
\label{eq:B}
  \Delta\vecp(e_0,T) & = & \matB(e_0,T) \vecq(T) \\
  \Rightarrow \quad 
  \covm^p_{\alpha\beta}(e_0,T) & = & \sum_{jk} B_{\alpha j}(e_0,T)
                                     B_{\beta k}(e_0,T) \, \covm^q_{jk}(T) 
\label{eq:covp}
\end{eqnarray}
which we would wish to average over the phase of the MBA orbit or mean anomaly $\ma$ at the time of observation $T.$

@@

\section{Impulse distribution}
\label{sec:impulse}
\subsection{Derivation}
Under the assumption that the distributions of masses and orbital
elements of MBAs are separable, the quantities needed to describe the
random walk of the asteroid's orbits can be expressed as an integral
over the absolute magnitude $H$ of the deflectors, the velocity $v$,
impact parameter $b,$ and polar/azimuthal angles $\theta,\phi$ of the
unit vector of the impulse direction $\bhat$:
\begin{eqnarray}
  \left\langle n I_\gamma^2 \right\rangle & = & \int
                                              dH\,dv\,db\,d\theta\,d\phi
                                             \frac{dN}{dH} \frac{dn}{dv\,db\,d\theta\,d\phi}
                                              I^2_\gamma(H,v,b,\theta\phi)
  \\
  & = & \int dH \frac{dN}{dH} \int dv
        \frac{dn}{dA\,dv\,d\theta\,d\phi} \hat b_\gamma^2(\theta,\phi)
        \int_{b_{\rm min}}^{b_{\rm max}} 2\pi b\,db\,
        \left(\frac{2\sigma_M(H)}{bv}\right)^2 \\
  & = & 8\pi \int dH \frac{dN}{dH} \sigma^2_M(H)  \int dv
        \frac{dn}{dA\,dv} v^{-2} \left\langle\hat b_\gamma^2(v)\right\rangle
        \log(b_{\rm max}/b_{\rm min}).
\end{eqnarray}
The second row adopts the (numerically verified) assumption that the
distribution of impact parameter $b$ will always be $\propto dA=2\pi
b\,db$ in the range of $b$ of interest to us.  The mean rate of encounters
between a deflector-tracer pair, as a function of the impact velocity
and geometry, is $dn/dA\,dv\,d\theta\,d\phi,$ which we will determine
through numerical measurements with the orbits of all known MBAs. The
$H$ distribution of MBAs is $dN/dH.$ We define $\sigma_M(H)$ via
\begin{equation}
  \sigma_M(H) \equiv \left\{\begin{array}{cl}
                              \sigma_M, & M(H) > \sigma_M \\
                              M(H),        & M(H) \le \sigma_M
                            \end{array} \right.
  \label{eq:sigmaMH}
\end{equation}
such that it represents the error in the ephemeris due to uncertainty
in the masses of MBAs included in the ephemeris (first row), and due
to the entirety of the masses of MBAs omitted from the ephemeris
(second row).

@@

\begin{figure}
  \centering
  \includegraphics[width=0.8\textwidth]{isq.pdf}
  \caption{Each panel shows statistics of the encounters between
    tracer MBAs from one region and deflector MBAs from another.  As a
    function of the relative velocity at impact, the three curves show
    the mean squared value of the radial, azimuthal, and vertical
    components of the impulse direction $\bhat.$  The horizontal
    dashed line at value $1/3$ would be the result for isotropic
    encounters, but we see that the azimuthal component of the impulse
    tends to be larger, with a variety of behaviors.  The gray regions
    trace the distribution $dN/dv$ of the rate of interactions per
    tracer per orbit per interval in $v.$  These curves have a common
    (arbitrary) normalization, so the relative heights properly
    reflect the frequency of encounters between different MBA
    regions.}
  \label{fig:rtz}
\end{figure}

@@

\begin{figure}
  \centering
  \includegraphics[width=0.8\textwidth]{mbacounts.pdf}
  \caption{Differential counts of MBAs cataloged by the MPC are
    plotted vs absolute magnitude $H$  
    for the Inner, Middle, Outer, and ``Other'' regions of
    the main belt, and for their total.  We assume that they are
    complete for $H\le17,$ and extrapolate to fainter sources using
    the functional form from \eqq{eq:lsstmba} shown as the dotted
    curve.}
  \label{fig:counts}
\end{figure}
We also require the $H$ distribution of potential deflectors.
Figure~\ref{fig:counts} plots our assumptions for $dN/dH.$  The MPC
catalog is likely close to complete in all three regions for $H\le17,$
so we will use the MPC counts directly in this regime.  Note that the
three regions have similar shapes of $dN/dH.$  For $H>17,$ we adopt
the functional form for the cumulative MBA counts given by
initial LSST predictions \citep{lsstbook}:
\begin{equation}
  N(<H) \propto \frac{10^{0.43(H-15.7)}}{10^{0.18(H-15.7)} +
    10^{-0.18(H-15.7)}},
\label{eq:lsstmba}
\end{equation}
normalizing this curve to the observed $H<17$ counts for each region.

Our integral for the uncertainty due to unmodelled MBA gravitation
requires the RMS error $\sigma_M(H)$ between the true mass $M$ and the
mass used (if any) in the ephemeris model.
The two largest asteroids, Ceres and Vesta, have masses determined to
high precision by the \textit{Dawn} spacecraft \citep{Ceres,Vesta}.  Thus despite holding
half of the $\approx 10^{-9} M_\odot$ mass of the asteroid belt, their $\sigma_M$
values\footnote{More precisely, the uncertainties in their $GM$ values
  relative to $GM_\odot$.} are 
$\le3\times10^{-15} M_\odot$ and
% Konilov papers: 1.2e-5 km^3/s^2 for Vesta, 4e-4 for Ceres
this unknown part of their contribution to other MBA's motion is 
unimportant.  Aside from a handful of other asteroids in
well-characterized binary systems or visited by spacecraft, most other knowledge of MBA masses
has and will come from mutual gravitational encounters.   An
individual asteroid's mass can be estimated from its measured effect  on the
few MBAs on which it imparts the largest impulses.  The accuracy
$\sigma_M$ of this determination will vary, depending upon the
geometry  and timing of
each deflector MBA's encounters, and the measurement errors on its tracers.  These
factors are independent of the mass of the deflector.  We will
assign a value $\sigma_M$ to the RMS uncertainty in MBA mass
attainable from following a few individual tracers, and calculate
the Brownian-motion noise as a function of $\sigma_M.$  We take
\begin{equation}
  \sigma_M(H) = \textrm{min}\left[ \sigma_M, M(H) \right],
\end{equation}
because a given MBA could have a mass determination accurate to
$\sigma_M$ if included in the ephemeris model, or if $M(H)<\sigma_M,$
it would be omitted from the ephemeris model.

@@

\subsubsection{Individually detectable impulses}
Examining the $\sigma_I$ distribution forecasted for LSST data on tracers (Figure~\ref{fig:sigmaI}), we find that for impulses occurring near the midpoint of the survey, $\approx80\%$ of all cases yield $\sigma_I<10^{-4}\,{\rm m}\,{\rm
  s}^{-1},$ some an order of magnitude lower.  We conservatively
assume that any encounter producing  an impulse $>5\times$ this value, $I_{\rm det}=5\times 10^{-4}\,{\rm m}\,{\rm
  s}^{-1},$ can be used to directly constrain the relevant deflector and include it in the ephemeris model.
When normalized by $v_c,$ we have $I_{\rm  det}=10^{-8}$ at 10\% accuracy across the main belt.

@@

\subsection{Element shifts from impulses}
\label{sec:elements}
The orbital shifts from an impulse can be derived from Gauss's
equations, \eg\ as presented by \citet[][Section 1.9.2]{tremaine}, but
we provide here a derivation customized to our case.
Five of our orbital elements $\vecq$ will be taken from constants of the motion.  The mean anomaly $\ma$ is not appropriate as the sixth, time-dependent element of $\vecq$ because the derivative $d\ma/dI$ can diverge as $e_0\rightarrow 0.$ Instead we introduce $\tau = \ma+\lop,$ which we will show does have a shift $\Delta\tau$ during an impulse that has finite and linear response to $\vecI$.  Between impulses, $\tau$ advances with the mean motion as $a^{-3/2}t.$  We define a coordinate system that is uniformly rotating with the unperturbed $\tau_0=t+\ma_0.$ The two smoothly rotating unit vectors $\uhat_\parallel=(\cos \tau_0, \sin \tau_0)$ and
$\uhat_\perp=(-\sin \tau_0, \cos \tau_0)$ satisfy $\uhat_\parallel
\times \uhat_\perp = \zhat,$ and we will use subscripts $\parallel$
and $\perp$ to represent projections onto these components.  In
particular, the components $(e_\parallel,e_\perp)$ of the ellipticity
vector $\vece$ have the values $e_\parallel=e\cos \ma$ and $e_\perp=-e\sin \ma$ in the unperturbed orbit.
The true anomaly $\nu,$ radius $r$, and azimuthal angle $\phi,$ and
the velocity components of the initial orbit are, to first order in $e$,
\begin{eqnarray}
  \ma & = & \ma_0 + t = \tau_0,  \nonumber \\
  \nu & = & \ma - 2 e_\perp \nonumber \\
  \phi & = & \nu + \lop = \tau - 2e_\perp, \nonumber\\
  r & = & 1-e_\parallel, \nonumber \\
  v_r & = & -e_\perp \nonumber \\
  v_t & = & 1 + e_\parallel
            \label{eq:kepler}
\end{eqnarray}

@@

Figure~\ref{fig:result} plots the resulting noise level $\sigma_\phi=\sqrt{\Var(\phi)}$ of the calculation for the Middle regions of MBAs, showing both the linear RMS displacements on the left axis and the angular displacements on the right axis.  Each point on the curve is the $\sigma_\phi$ attributable to deflector MBAs that are smaller (fainter) than the $H$ value on the lower axis, or the mass on the top axis.  Thus the left-most point on each curve marks the total noise integrated over the full population of deflectors. Different color lines show the results for the values of $N$ and $\sigma_M$ as marked astride each line.  Notable features of these result are:
\begin{itemize}
\item We find that \emph{the linear displacements are nearly identical for Inner, Middle, and Outer regions of the asteroid belt.}  The angular displacements vary slightly, inversely to their distance from the Sun.
\item We plot the azimuthal noise.  \emph{The radial noise results look almost identical, but $\approx 7\times$ lower RMS, and likewise the vertical Brownian motion is  very similar but scaled $\approx45\times$ lower than the azimuthal RMS.}
\item For longer time periods $T$, the azimuthal and radial noise will scale as
$T^{3/2},$ the vertical
noise as $T^{1/2}.$
\item The largest contributions to the unmodelled Brownian motion are
made by deflectors with mass and $H$ at the ``shoulder'' of each
curve, which will also correspond to the MBAs with masses $M\sim
\sigma_M$.  Above this mass, the noise per deflector is constant but
the deflectors become scarce at larger $M.$ Below this mass, $dN/dH$
is rising to fainter/smaller bodies, but the factor $M^2(H) dN/dH$
that determines the Brownian motion is dropping.
\end{itemize}

>>>>

\newcommand{\Var}{\textrm{Var}}
\newcommand{\vcirc}{v_c}
\newcommand{\Msun}{M_\odot}
\newcommand{\covm}{C}
\newcommand{\lop}{\varpi}   % longitude of perihelion
\newcommand{\ma}{\ell}  % Mean anomaly

@@

High-precision measurements of positions of Solar System bodies have
great value to multiple scientific pursuits.  They can reveal
gravitational accelerations due to undiscovered bodies such as a
hypothetical ``Planet X'' \citep[\eg][]{holmanP9,inpopP9,trojans,occultations} or passing primordial
black holes \citep[\eg][]{pbh}, determine
the masses of known bodies \citep[\eg][]{goffin,baer}, test the laws
of physics \citep[\eg][]{inpopgraviton}, and characterize 
the non-gravitational forces that drive many forms of planetary
migration, such as the maintainence of the near-Earth asteroid (NEA)
population \citep[reviewed by][]{yarkovsky}.  While most applications of precision orbit determination
have used the 8 major planets as test particles, there are now $>10^6$
known minor planets available for this task.  These small bodies' astrometric
information content is growing extremely rapidly through large-scale
survey projects that can produce
milliarcsecond-scale accuracy for magnitude-limited small-body populations, such as Gaia \citep{gaiass3,gaiafpr},
PanSTARRS,\footnote{\url{https://www2.ifa.hawaii.edu/research/Pan-STARRS.shtml}}
and particularly the upcoming Rubin observatory
\textit{Legacy Survey of Space and Time} (LSST).\footnote{\url{https://rubinobservatory.org/}}  The great majority
of currently-tracked objects are main-belt asteroids (MBAs), so we
wish to investigate the potential power of MBAs as gravitational test
bodies.

@@

In principal, one can collect all of the positional information for
all $>10^6$ known solar-system bodies, and fit them to a model with
a free state vector and mass for each body.  Calculating the influence
of $>7\times10^6$ free parameters on $10^8$ or so observables is,
however, unlikely to be computationally feasible, and most of the
bodies' masses will be lost in measurement noise and/or be degenerate
with other bodies' masses.  The most ambitious attempt at a global
solution that we are aware of is that of \citet{goffin}, who was able
to include 250 MBAs as gravitating bodies in the fit to the full
corpus of observations at the time.  For only $\approx 1/2$ of these
bodies did the fit yield masses $>3\times$ the uncertainties.  With
vast improvements in data and computing power since then and in the
next decade, the number of detectable
asteroid masses should increase, but there will always be
uncertainties that accrue from the unmodelled deflectors and the
finite $\sigma_M$ of modelled deflectors.  These remnant uncertainties
will in turn serve as noise in attempts to model new sources of
acceleration.

@@

\section{Calculation overview}
We aim to calculate the Brownian motion variance to an accuracy of a
factor $\approx 2.$
Greater accuracy in the calculation is not warranted
since some of the principal inputs have substantial uncertainty.  The
number of MBAs vs mass is 
uncertain because of unknowns in the conversion
from $H$ to mass; the status of future observations and $\sigma_M$ is
also unclear.  We are hence justified in retaining only the
contributions to our results that are leading
order in the orbital eccentricity $e$ of the tracer MBA.
We will assume throughout that the joint
distribution of mass (or $H$) and orbital elements is separable into a
mass distribution and an orbital distribution.  In this scenario, the
Brownian motion is independent of the tracer mass, and involves an
integral over the mass distribution of the deflecting bodies.

We adopt a cylindrical coordinate system $(r,\phi,z),$ with $\zhat$
normal to the initial orbital plane, and $\phi=z=0$ toward the
perhelion of the initial orbit, \ie\ both the initial ascending node
$\Omega_0$ and longitude of perihelion $\lop_0$ equal to zero.  The
unit vectors $\rhat, \phat$ rotate with the target asteroid.  The mean
anomaly at time $t=0$ is $\ma_0.$  In general, subscripts of $0$ will
indicate properties of the unperturbed orbit.

@@

We describe all encounters with other asteroids in the impulse approximation, defining $\vecI$ as the $\Delta\vecv$ imparted on the target by the deflector.
In our units, the gravitational impulse imparted by a deflector of mass $M_d$ approaching at impact parameter $\vecb=b\,\bhat$ and relative velocity $v$ is
\begin{equation}
  \vecI = 2 \frac{M_d}{\Msun} (bv)^{-1} \bhat.
  \label{eq:impulse}
\end{equation}
All of our results will be derived at leading order in $\vecI,$ which
is very well justified by the small size of MBA-induced impulses.

@@

From this knowledge of the impulse distribution, our goal is to obtain
the covariance matrix of the deviations of the target's position from
the initial orbit, after some time $T.$ The position observables are
the range, plus the celestial latitude and longitude of the target.  We will simplify our results by assuming a heliocentric observer, so that the observational position vector is $\vecp\equiv (r,\phi,\theta=z/r).$ The quantity we seek is the covariance matrix $\covm^p$ of the observations attributable to the accumulated gravitational perturbations:
\begin{equation}
  \covm^p  \equiv  \left\langle  \Delta\vecp \Delta\vecp^T\right\rangle,  \label{eq:Cp}
\end{equation}
where the angle brackets indicate an average over possible
realizations of the impulse history, and over the mean anomaly $\ma$
at the time of observation.  To do so, we will introduce an
intermediate set of 6 orbital elements $\vecq,$ selected to respond linearly to $\vecI$ at $|I|\ll 1.$
We will derive the matrix $\matA$ that describes the orbital-element shifts at time $T$ that arise from an impulse at time $t_i$: 
\begin{equation}
  \Delta \vecq(e_0,T,t_i) = \matA(e_0, T, t_i) \, \vecI.
  \label{eq:A}
\end{equation}
In our first-order perturbation theory, $\Delta\vecq(T)$ will be the sum of the $\Delta\vecq_i$ imparted by all impulses applied at times $0<t_i<T.$  Because the impulses are uncorrelated, $\vecq$ will therefore be the result of a random walk.  The distribution will have a covariance matrix $\covm^q(T) \equiv \left\langle \Delta\vecq(T) \Delta\vecq^T(T) \right\rangle$ whose elements are
\begin{eqnarray}
  \covm^q_{jk}(e_o,T) & = & \left\langle \sum_{i,\gamma} \matA_{j\gamma}(e_0,T,t_i) \matA_{k\gamma}(e_0,T,t_i) I^2_{i,\gamma} \right\rangle \\
           & = & \sum_\gamma \int dt_i \matA_{j\gamma}(e_0,T,t_i) \matA_{k\gamma}(e_0,T,t_i) \left\langle n I_\gamma^2\right\rangle.
\label{eq:Cqjk}
\end{eqnarray}
In the first line, the sum $i$ runs over the impulses and $\gamma$ runs over the components $r,\phi,z$ of the impulse. The second line evaluates the expectation value of averaging over realizations of the random walk of impulses, exploiting the independence of the individual impulses from each other.

The last element of our calculation will be a conversion from the
element shifts $\Delta\vecq$ into the observed displacements $\Delta\vecp$
at time $T.$  In linear perturbation theory this will again be expressible as a matrix
\begin{eqnarray}
\label{eq:B}
  \Delta\vecp(e_0,T) & = & \matB(e_0,T) \vecq(T) \\
  \Rightarrow \quad 
  \covm^p_{\alpha\beta}(e_0,T) & = & \sum_{jk} B_{\alpha j}(e_0,T)
                                     B_{\beta k}(e_0,T) \, \covm^q_{jk}(T) 
\label{eq:covp}
\end{eqnarray}
which we will average over the phase of the MBA orbit at the
observation time $T,$ \ie\ average over mean anomaly $\ma.$

@@

\section{Impulse distribution}
\label{sec:impulse}
\subsection{Derivation}
Under the assumption that the distributions of masses and orbital
elements of MBAs are separable, the quantities needed to describe the
random walk of the asteroid's orbits can be expressed as an integral
over the absolute magnitude $H$ of the deflectors, the relative velocity $v$,
impact parameter $b,$ and polar/azimuthal angles $\theta,\phi$ of the
unit vector of the impulse direction $\bhat$:
\begin{eqnarray}
  \left\langle n I_\gamma^2 \right\rangle & = & \int
                                              dH\,dv\,db\,d\theta\,d\phi
                                             \frac{dN}{dH} \frac{dn}{dv\,db\,d\theta\,d\phi}
                                              I^2_\gamma(H,v,b,\theta,\phi)
  \\
  & = & \int dH \frac{dN}{dH} \int dv
        \frac{dn}{dA\,dv\,d\theta\,d\phi} \hat b_\gamma^2(\theta,\phi)
        \int_{b_{\rm min}}^{b_{\rm max}} 2\pi b\,db\,
        \left(\frac{2\sigma_M(H)}{bv}\right)^2 \\
  & = & 8\pi \int dH \frac{dN}{dH} \sigma^2_M(H)  \int dv
        \frac{dn}{dA\,dv} v^{-2} \left\langle\hat b_\gamma^2(v)\right\rangle
        \log(b_{\rm max}/b_{\rm min}).
\end{eqnarray}
The second row adopts the (numerically verified) assumption that the
distribution of impact parameter $b$ will always be $\propto dA=2\pi
b\,db$ in the range of $b$ of interest to us.  The mean rate of encounters
between a deflector-tracer pair, as a function of the impact velocity
and geometry, is $dn/dA\,dv\,d\theta\,d\phi,$ which we will determine
through numerical measurements with the orbits of all known MBAs. The
$H$ distribution of MBAs is $dN/dH.$ We define $\sigma_M(H)$ via
\begin{equation}
  \sigma_M(H) \equiv \left\{\begin{array}{cl}
                              \sigma_M, & M(H) > \sigma_M \\
                              M(H),        & M(H) \le \sigma_M
                            \end{array} \right.
  \label{eq:sigmaMH}
\end{equation}
such that it represents the error in the ephemeris due to uncertainty
in the masses of MBAs included in the ephemeris (first row), and due
to the entirety of the masses of MBAs omitted from the ephemeris
(second row).

@@

\begin{figure}
  \centering
  \includegraphics[width=0.8\textwidth]{isq.pdf}
  \caption{Each panel shows statistics of the encounters between
    tracer MBAs from one region and deflector MBAs from another.  As a
    function of the relative velocity at impact, the three curves show
    the mean squared value of the radial, azimuthal, and vertical
    components of the impulse direction $\bhat.$  The horizontal
    dashed line at value $1/3$ would be the result for isotropic
    encounters, but we see that the azimuthal component of the impulse
    tends to be larger, with a variety of behaviors.  The gray regions
    trace the distribution $dN/dv$ of the rate of interactions per
    tracer per orbit per interval in $v.$  These curves have a common
    (arbitrary) normalization, so the relative heights properly
    reflect the relative frequency of encounters among
    pairs of deflector and tracer regions.}
  \label{fig:rtz}
\end{figure}

@@

\begin{figure}
  \centering
  \includegraphics[width=0.8\textwidth]{mbacounts.pdf}
  \caption{Differential counts of MBAs cataloged by the MPC are
    plotted vs absolute magnitude $H$  
    for the Inner, Middle, Outer, and ``Other'' regions of
    the main belt, and for their total.  We assume that they are
    complete for $H\le17,$ and extrapolate to fainter sources using
    the functional form from \eqq{eq:lsstmba} shown as the dotted
    curve. The Inner and Outer regions are slightly rescaled to help
    visual comparison of the curves' shapes in the three MBA regions.}
  \label{fig:counts}
\end{figure}
We also require the $H$ distribution of potential deflectors.
Figure~\ref{fig:counts} plots our assumptions for $dN/dH.$  The MPC
catalog is likely close to complete in all three regions for $H\le17,$
so we will use the MPC counts directly in this regime.  Note that the
three regions have similar shapes of $dN/dH.$  For $H>17,$ we adopt
the functional form for the cumulative MBA counts given by
initial LSST predictions \citep{lsstbook}:
\begin{equation}
  N(<H) \propto \frac{10^{0.43(H-15.7)}}{10^{0.18(H-15.7)} +
    10^{-0.18(H-15.7)}},
\label{eq:lsstmba}
\end{equation}
normalizing this curve to the observed $H<17$ counts for each region.

Our integral for the uncertainty due to unmodelled MBA gravitation
requires the RMS error $\sigma_M(H)$ between the true mass $M$ and the
mass used (if any) in the ephemeris model.
The two largest asteroids, Ceres and Vesta, have masses determined to
high precision by the \textit{Dawn} spacecraft \citep{Ceres,Vesta}.  Thus despite holding
half of the $\approx 10^{-9} M_\odot$ mass of the asteroid belt, their $\sigma_M$
values\footnote{More precisely, the uncertainties in their $GM$ values
  relative to $GM_\odot$.} are 
$\le3\times10^{-15} M_\odot$ and
% Konilov papers: 1.2e-5 km^3/s^2 for Vesta, 4e-4 for Ceres
this unknown part of their contribution to other MBA's motion is 
unimportant.  Aside from a handful of other asteroids in
well-characterized binary systems or visited by spacecraft, most other knowledge of MBA masses
has and will come from mutual gravitational encounters.   An
individual asteroid's mass can be estimated from its measured effect  on the
few MBAs on which it imparts the largest impulses.  The accuracy
$\sigma_M$ of this determination will vary, depending upon the
geometry  and timing of
each deflector MBA's encounters, and the measurement errors on its tracers.  These
factors are independent of the mass of the deflector.  We will
assign a value $\sigma_M$ to the RMS uncertainty in MBA mass
attainable from following a few individual tracers, and calculate
the Brownian-motion noise as a function of $\sigma_M.$

@@

\subsubsection{Individually detectable impulses}
Examining the $\sigma_I$ distribution forecasted for LSST data on
tracers by \citet{negin}, we find that for impulses occurring near the midpoint of the survey, $\approx80\%$ of all cases yield $\sigma_I<10^{-4}\,{\rm m}\,{\rm
  s}^{-1},$ some an order of magnitude lower.  We conservatively
assume that any encounter producing  an impulse $>5\times$ this value, $I_{\rm det}=5\times 10^{-4}\,{\rm m}\,{\rm
  s}^{-1},$ can be used to directly constrain the relevant deflector and include it in the ephemeris model.
When normalized by $v_c,$ we have $I_{\rm  det}=10^{-8}$ at 10\% accuracy across the main belt.

@@

\subsection{Element shifts from impulses}
\label{sec:elements}
The orbital shifts from an impulse can be derived from Gauss's
equations, \eg\ as presented by \citet[][Section 1.9.2]{tremaine}, but
we provide here a derivation customized to our case.
Five of our orbital elements $\vecq$ will be taken from constants of the motion.  The mean anomaly $\ma$ is not appropriate as the sixth, time-dependent element of $\vecq$ because the derivative $d\ma/dI$ can diverge as $e_0\rightarrow 0.$ Instead we introduce $\tau = \ma+\lop,$ which we will show does have a shift $\Delta\tau$ during an impulse that has finite and linear response to $\vecI$.  Between impulses, $\tau$ advances with the mean motion as $a^{-3/2}t.$  We define a coordinate system that is uniformly rotating with the unperturbed $\tau_0=t+\ma_0.$ The two smoothly rotating unit vectors $\uhat_\parallel=(\cos \tau_0, \sin \tau_0)$ and
$\uhat_\perp=(-\sin \tau_0, \cos \tau_0)$ satisfy $\uhat_\parallel
\times \uhat_\perp = \zhat,$ and we will use subscripts $\parallel$
and $\perp$ to represent projections onto these components.  In
particular, the components $(e_\parallel,e_\perp)$ of the ellipticity
vector $\vece$ have the values $e_\parallel=e\cos \ma$ and $e_\perp=-e\sin \ma$ in the unperturbed orbit.
The true anomaly $\nu,$ radius $r$, and azimuthal angle $\phi,$ and
the velocity components of the unperturbed orbit are, to first order in $e$,
\begin{eqnarray}
  \ma & = & \ma_0 + t = \tau_0,  \nonumber \\
  \nu & = & \ma - 2 e_\perp \nonumber \\
  \phi & = & \nu + \lop = \tau - 2e_\perp, \nonumber\\
  r & = & 1-e_\parallel, \nonumber \\
  v_r & = & -e_\perp \nonumber \\
  v_t & = & 1 + e_\parallel
            \label{eq:kepler}
\end{eqnarray}

@@

Figure~\ref{fig:result} plots the resulting noise level $\sigma_\phi=\sqrt{\Var(\phi)}$ of the calculation for the Middle regions of MBAs, showing both the linear RMS displacements on the left axis and the angular displacements on the right axis.  Each point on the curve is the $\sigma_\phi$ attributable to deflector MBAs that are smaller (fainter) than the $H$ value on the lower axis, or the mass on the top axis.  Thus the left-most point on each curve marks the total noise integrated over the full population of deflectors. Different color lines show the results for the values of $N$ and $\sigma_M$ as marked astride each line.  Notable features of these result are:
\begin{itemize}
\item \emph{The linear displacements are nearly identical for Inner,
    Middle, and Outer regions of the asteroid belt.}  The angular
  displacements vary slightly, inversely to their distance from the
  Sun.  Figure~\ref{fig:result} plots only the Middle MBAs'
  deviations, although the count $N$ of modelled MBAs is the sum for
  all three regions.
\item We plot the azimuthal noise.  \emph{The radial noise results
    have nearly identical scaling with $H$ and $\sigma_M$, but
    $\approx 7\times$ lower RMS. Likewise the vertical Brownian motion
    has the same shape but $\approx45\times$ lower than the azimuthal RMS.}
\item For longer time periods $T$, the azimuthal and radial noise will scale as
$T^{3/2},$ the vertical
noise as $T^{1/2}.$
\item The largest contributions to the unmodelled Brownian motion are
made by deflectors with mass and $H$ at the ``shoulder'' of each
curve, which will also correspond to the MBAs with masses $M\sim
\sigma_M$.  Above this mass, the noise per deflector is constant but
the deflectors become scarce at larger $M.$ Below this mass, $dN/dH$
is rising to fainter/smaller bodies, but the factor $M^2(H) dN/dH$
that determines the Brownian motion is dropping.
\end{itemize}
